---
title: "ArcticDEM - Geomorphometrical extraction of Eskers"
author: "Julian Stuber"
subtitle: subtitle1
output:
  html_document:
    code_folding: show
    fig_caption: yes
    highlight: tango
    number_sections: no
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '2'
---

<style>
body {
text-align: justify}
</style>


# Info & Setup
The focus of this data challenge will be on the methods for point pattern analysis introduced in Sessions 2 to 5, using a data set of crime incidents in several wards of the city center of York, UK.
The following data sets are available for the assignment at hand:

`York_crimes_2016-09.csv`: contains observed crimes in the city center of York.

`York_gastronomy.shp`: containing bars, restaurants and pubs located in the city center of York, extracted from OpenStreetMap.

`York_wards_city_center.shp`: contains the boundaries of the wards of the city center of York.

`York_population_density.tif`, an interpolated rasterlayer of the population density of York

Small annotation: Hopefully the internal links work. Sometimes they do and sometimes they do not. 

## *Clean project, set options*
```{r}
knitr::opts_chunk$set(echo = TRUE)
#rm(list=ls())             # Clean up the environment
# To start with a clean slate, use the interactive "Session > Restart R" and
# "Session > Clear Workspace ..." menu options in RStudio, rather than rm(list=ls())
# For reasons why not to use rm(list=ls()), see here:
# https://www.tidyverse.org/articles/2017/12/workflow-vs-script/

# set directories
directory    <- getwd()
dataFolder   <- file.path(directory, "data")
RFolder      <- file.path(directory)
outputFolder <- file.path(directory, "output")

# Save plot/ map as .pdf to figures folder
saveFigure <- function(plotname, filename){
  pdf(paste("figures/", filename, ".pdf", sep = ""))
  print(plotname)
  dev.off()
}

options(scipen=6)         # Display digits, not the scientific version
options(digits.secs=6)    # Use milliseconds in Date/Time data types
options(warning=FALSE)    # Don't show warnings
par(mfrow=c(1,1))         # Reset plot placement to normal 1 by 1
```


## *Libaries*
Chunk checking for missing packages from existing list based on precedent lab sessions. 
```{r include = FALSE}
## Default repository
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org" 
       options(repos=r)
})

pkgTest <- function(x)
  {
    if (!require(x,character.only = TRUE))
    {
      install.packages(x,dep=TRUE)
        if(!require(x,character.only = TRUE)) stop("Package not found")
    }
}

# Package test
pkgTest("raster")
pkgTest("rgdal")
pkgTest("stringr")
pkgTest("sf")
pkgTest("sp")
```

## *CRS strings*
```{r, warning=FALSE, message=FALSE, fig.align="center"}
# Projection strings for the CRS needed for this project:
CRS3413   = "+init=epsg:3413"   # 
CRS9822   = "+init=epsg:9822"   # 
CRS4326   = "+init=epsg:4326"   # WGS84 Grid
# CRS102001 = "+init=epsg:102001" #
# CRS102002 = "+init=epsg:102002" # 
crs_wgs84 <- st_crs(4326)
```

## *Loading data from net, create sub-diretory, unpack (.tar.gz)*
```{r, warning=FALSE, message=FALSE}
# get URL of desired mosaic file, paste it here when prompted:
# "https://data.pgc.umn.edu/elev/dem/setsm/ArcticDEM/mosaic/v3.0/32m/26_12/26_12_32m_v3.0.tar.gz" #as an example URL
fn_https <- readline(prompt = "Paste URL: ")
# prepare string for dir.create in next step
fn_str <- stringr::str_sub(fn_https, -21, -13)
fn_local <- file.path(dataFolder, fn_str)
# create sub-directory in data folder for specific tile and resolution
ifelse(!dir.exists(fn_local), dir.create(fn_local, recursive = T), "Folder exists already") 

print("Your file is being downloaded...")
# pull .tar.gz file from internet server
download.file(fn_https, 
              destfile = file.path(fn_local, "tmp.tar.gz") , mode='wb')

untar(file.path(fn_local, "tmp.tar.gz"), list = T) #check contents
untar(file.path(fn_local, "tmp.tar.gz"), exdir = fn_local) # if exdir not specified, new data will appear in wd

file.remove(file.path(fn_local, "tmp.tar.gz"))

# store location of eskers_Canada.shp
vn_https <- "https://www.researchgate.net/profile/Rob-Storrar/publication/284715993_Canadian_eskers_shapefile/data/5656d22208ae4988a7b505f7/Storrar-et-al-2013-JoM.zip"
# pull .zip file from internet server
download.file(vn_https, 
              destfile = file.path(dataFolder, "eskers_Canada.zip") , mode='wb')
unzip(file.path(dataFolder, "eskers_Canada.zip"), exdir = dataFolder)
file.remove(file.path(dataFolder, "eskers_Canada.zip"))
```


## *Open downloaded raster file, crop tiles into 5000 x 5000cells*
```{r, warning=FALSE, message=FALSE}
# get file name of only(!) .tif-file in tile-folder
fnames <- list.files(fn_local, pattern = "\\.tif$") #c heck that no more than one .tif file inside folder!
# open DEM inforamtion in rstudio
demtile <- raster(file.path(fn_local, fnames))

# quick plot
# plot(demtile, col = rev(terrain.colors(50)))

# check CRS and reference ellipsoid
crs(demtile)
proj4string(demtile) <- CRS3413
crs(demtile)
```

## *Open downloaded vector file, crop to raster extent*
```{r, warning=FALSE, message=FALSE}
# open shapefile of eskers in rstudio
eskersca <- readOGR(dsn = dataFolder, layer = "eskers_Canada")
eskers <- st_read(system.file("vectors/cities.shp", package = "rgdal"))

# check CRS and reference ellipsoid
crs(eskersca)
proj4string(eskersca) <- CRS3413
crs(eskersca)

# quick plot
plot(eskersca, col = rev(terrain.colors(50)))
```